---
title: "risk-perception"
author: "Nick Gauthier"
date: "9/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(zoo)
```




```{r}
test <- tibble(agent = 1:5,
       alpha = 1.5, #runif(5),
       beta = 1.5) %>%
  mutate(mu = alpha / (alpha + beta))

test2 <- c((runif(25) < .2),(runif(25) < .5), runif(25) < .1)
tibble(time = 1:75,
       drought = test2) %>%
  mutate(period = round(time, -1)) %>%
  add_count(period) %>%
  filter(period >= 10) %>%
  group_by(period) %>%
  summarise(drought = sum(drought)) %>%
  ggplot(aes(period, drought)) +
  geom_col()
```

```{r}
learn <- function(agents, event) {
  agents %>%
    mutate(alpha = event + alpha,
           beta = beta + 1 - event,
           mu = alpha / (alpha + beta),
           estimate = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
}

accumulate(test2, ~learn(.x, .y), .init = test) %>% bind_rows(.id = 'time') %>% qplot(as.numeric(time), estimate, data = ., geom = 'point')
```

```{r}
learn(test, TRUE)
```


The basic risk assessment agorithm. Let's look at the risk of a drought year being 0.2. In a wet year you get 3 wheat and 2 barley, and in a dry year you get 0 wheat and 1 barley. What does that look like?

```{r}
wheat_dry <- 0
wheat_wet <- 1.6
barley_dry <- 0.93
barley_wet <- 1.18

calc_yield <- function(wheat_prop, risk) {
  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
  barley + wheat
}


util_wet <- wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)
util_dry <- wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)

util_wet * (1 - risk)

calc_yield(.92, .5)

calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)

calc_yield(1, .2)

plot(calc_yield((0:10)/10, .2))

expand_grid(wheat_prop = (0:10)/10,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)


expand_grid(wheat_prop = (0:10)/1,
       risk = (0:10)/10) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop))) %>%
  gather(type, util, util_wet:util_dry) %>%
  ggplot(aes(risk, util, color = type)) + geom_line() + facet_wrap(~wheat_prop)


calc_prop <- function(risk) {
  pb <- barley_wet * (1 - risk) - barley_dry * risk
  pw <- wheat_wet * (1 - risk) - wheat_dry * risk

  pb / (pb + pw) 
}

plot(calc_prop((0:10)/10))
```


```{r}
calcNE <- function(risk){
  (118 - 25 * risk) / (278 - 185 * risk)
}


(barley_wet - .2 * (barley_dry + wheat_dry + wheat_wet + barley_wet)) / (barley_wet + wheat_wet)

calcNE(1)


plot(1.6 * (1:10)/10 + 1.18 * (10:1)/10)
```


so we're imagining a yield of 1 for barley and 2 for wheat, with equal prportions of each. then in the next time step we'd revise the proportion to take away .25
```{r}
reallocate <- function(wheat_prop, risk) {
  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
}

accumulate(1:10, ~reallocate(.x), .init = .4) %>% plot

# so this works as long as you're between 0 and 1
```


```{r}

prior <- function(m,n){
a = n * m
b = n * (1 - m)
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}

prior(.5, 2) %>% plot
```

```{r}
likelihood <- function(N,Y){
a <- Y + 1
b <- N - Y + 1
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}

likelihood(1, 1)%>% plot
```

```{r}
posterior <- function(m,n,N,Y){
a <- Y + (n*m) -1
b <- N - Y + (n*(1-m)) - 1
dom <- seq(0,1,0.005)
val <- dbeta(dom,a,b)
return(data.frame('x'=dom, 'y'=val))
}


-1
1 - 0 - 1

dbeta(seq(0,1,0.005), -1, 0)
prior(.5, 2) %>% plot
posterior(.5, 3, 1, 1) %>% plot
```


so now we get the same yields, but the proportions are different
```{r}
(1 *  0.625 - 2 * .375) * .375 * .625
```

```{r}
1- .375
```


```{r}
-1 * 0.0625 * (1 - 0.0625)
```



```{r}
yield_max_barley <- 2500
yield_max_wheat <- 3500
kcal_barley <- 3000
kcal_wheat <- 3540

calc_yield <- function(rainfall, crop = 'wheat') {
  if (crop == 'wheat') {
    yield <- yield_max_wheat * kcal_wheat * pmax(0, 0.51 * log(rainfall) + 1.03)  # annual rainfall impact on yields
  }
  
  if(crop == 'barley') {
    yield <- yield_max_barley * kcal_barley * pmax(0, 0.48 * log(rainfall) + 1.51) 
  }
  
  return(yield)
}


tibble(rainfall = (0:200) / 100,
       barley = calc_yield(rainfall, 'barley'),
       wheat = calc_yield(rainfall, 'wheat')) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + 
  geom_line() +
  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
  geom_hline(yintercept = yield_max_barley * kcal_barley)
```


```{r}
# generate some forgetting curves
forgetting_curves <- tibble(x = c(.01, .1, 1, 0)) %>%
  mutate(memory = map(x, ~exp(-1 * . *1:50))) %>%
  unnest(cols= c(memory)) %>%
  group_by(x) %>%
  mutate(time = 1:n())

ggplot(forgetting_curves, aes(time, memory)) +
  geom_line(aes(color = as.factor(x)), size = 1.2) +
  labs(title = 'Forgetting curves', x = 'Years', y = 'Relative memory strength\n') +
  scale_color_viridis_d(name = 'Decay rate') +
  theme_classic()
```

```{r}
farm <- function(households){
  households %>%
    mutate(yield = climatic_yield, #* n_inhabitants ^ labor_elasticity,
           yield_memory = yield, #map2(yield_memory, yield, remember),
           harvest = land * yield * .5 - land * sowing_rate) #%>%  # *.5 is for fallow
  #select(-yield)
}


remember <- function(yield_memory, yield){
  # rnorm(1, yield, yield * 0.0333) %>%  #memory is fuzzy
  append(yield_memory[-length(yield_memory)], yield, after = 0) # remove the last entry in the vector and add new yield to the begining
}


#Agents use the peak-end rule when accessing memory.
peak_end <- function(x){
  map_dbl(x, ~mean(c(.x[1], min(.x))))
}

```




```{r}
obs.na <- read.csv('salihli_p.csv')[10:620, 5]  / 10  # raw data in mm*10. subset time series as water years (i.e. first observation is october, last is september)

get.winter <- function(x){seq(0, length(x) - 12, 12) %>%
    sapply(function(y) y + 1:6) %>%
    as.numeric %>% extract(x, .)}# function to get Oct-Mar values from
obs<- obs.na %>% 
  rev %>% 
  na.locf %>% 
  rev %>% 
  `+`(na.locf(obs.na)) %>% 
  `/`(2) %>%  # fill NA with avg of adjacent values
  get.winter

# Now import TraCE-21k GCM data. Extract a time series for the grid cell covering the Salihli weather station. Add the model-output convective and large-scale precipitation variables to get total precipitation. Convert from m/s to mm/month. Subset time series to the calibration period (10 BP - AD 1940) and prediction period (10000 BP - 11 BP).

salihli.pt <- c(28.14, 38.48) %>% matrix(nrow = 1) %>% SpatialPoints  # create a point in Salihli to extract data from GCM

pc <- list.files('GCM Data/PRECC', full.names = T) %>%  # get a list of all GCM output files
  lapply(brick) %>%                                 # import GCM files as raster bricks
  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
  unlist                                            # combine into a single vector

pl <- list.files('GCM Data/PRECL', full.names = T) %>%  # get a list of all GCM output files
  lapply(brick) %>%                                 # import GCM files as raster bricks
  lapply(., raster::extract,salihli.pt) %>%                  # extract series from the grid cell containing Salihli
  unlist                                            # combine into a single vector

pt <- pc %>% add(pl) %>% multiply_by(60 * 60 * 24 * 30 * 1000)  # add precip components and convert from m/s to mm/month

cal <- pt %>% extract(122278:122877) %>% get.winter

pred <- pt %>% extract(2398:122397) %>% get.winter

pred.byfifty <- pred %>% seq_along %>% divide_by(300) %>% ceiling %>% split(pred,.)

calc.cdf <- function(x){
  pred <- x
  ct <- CDFt(obs, cal, pred)
  ct$DS
}
monthly.recon <- lapply(pred.byfifty, calc.cdf) %>% unlist
annual.recon <- monthly.recon %>% seq_along %>% divide_by(6) %>% ceiling %>% split(monthly.recon,.) %>%
  lapply(sum) %>% unlist %>% data.frame(date = seq(-10000,-1,1), p = ., color = cut(., c(0,200,300, 800)))

```

